---
title: "DBSCAN"
author: "Filippo De Boni"
date: "2026-04-08"
output: html_document
---

# DBSCAN

Cluster analysis is performed with a multimethod approach through partitional, hierarchical and density-based clustering methods.

DBSCAN is implemented in an unconventional way by imposing a k parameter. This approach is partially justified by the intent to compare the calculated clusterings with the ground truths, rather than the establishment of the real k parameter. In the original study the selection method (implemented with the Starczewski et al. 2020's approach) was pursued with human supervision. An automated approach is presented here, aiming at the balancing of some trade off: noise reduction, balance between clusters dimensions and k conformity with the ground truths. Results still appear underwhelming, with many clusterings showing too much noise (at least on this synthetic data); some work is in progress to make this pipeline work properly

```{r}
rm(list=ls())

library(readr)
library(ggfortify)
library(fpc)
library(clusterCrit)
library(cluster)
library(dbscan)
source("db.ground.truth.R")

agg1 <- read_csv("agg1.csv", col_types = cols(...1 = col_skip()))
agg1$Shape=factor(agg1$Shape)
agg1$Layer=factor(agg1$Layer)
agg1$Animal=factor(agg1$Animal)
agg1$ID=factor(agg1$ID)
agg1$Image=factor(agg1$Image)

set.seed(111)
```

Naming all the aggregated measures and marking them by sample size:

-   N\<=5: \*\*\* (In the original study the measures obtained by too few observations were deleted)

-   5\<N\<=10: \*\*

-   10\<N\<=20: \*

-   20\<N\<30: .

```{r}
Name=c(rep(NA,nrow(agg1)))
for (i in 1:nrow(agg1)){
  Name[i]=paste(agg1$Animal[i],"_",agg1$Image[i],"_",agg1$Shape[i],"_",agg1$Layer[i])
}
w=which(agg1$N<=5)
for (i in 1:length(w)){
  Name[w[i]]=paste(Name[w[i]],"***")
}
w=which(agg1$N>5 & agg1$N<=10)
for (i in 1:length(w)){
  Name[w[i]]=paste(Name[w[i]],"**")
}
w=which(agg1$N>10 & agg1$N<=20)
for (i in 1:length(w)){
  Name[w[i]]=paste(Name[w[i]],"*")
}
w=which(agg1$N>20 & agg1$N<30)
for (i in 1:length(w)){
  Name[w[i]]=paste(Name[w[i]],".")
}
```

Defining some classifications of interest:

-   **Cetacean**: do the observations come from a Cetacean's visual cortex?

-   **Aquatic**: the observations belong to a Terrestrial, an Aquatic or a Hippopotamus amphibius mammal's visual cortex?

-   **Diet/Binocularity**: which kind of diet/binocularity characterizes the observations' original subject?

-   **Order**: to which phylogenetic order does the subject the observations come from belong?

```{r}
w=c(7,11:13)
agg1=agg1[,-w]
acq=c(rep(0,270),rep(1,180),rep(0,60),rep(2,60),rep(0,120))
agg1=cbind(acq,agg1)
colnames(agg1)=c("Aquatic",colnames(agg1)[2:17])
agg1$Aquatic=factor(agg1$Aquatic,levels=c(0,1,2),labels=c("Terrestrial","Aquatic","Hippo"))
cet=c(rep(0,270),rep(1,180),rep(0,240))
agg1=cbind(cet,agg1)
colnames(agg1)=c("Cetacean",colnames(agg1)[2:18])
agg1$Cetacean=factor(agg1$Cetacean,levels=c(0,1),labels=c("No","Yes"))
w1=which(agg1$Animal=="Tursiops truncatus" | agg1$Animal=="Grampus griseus" | agg1$Animal=="Physeter macrocephalus" | agg1$Animal=="Globicephala melas")
w2=which(agg1$Animal=="Balaenoptera physalus")
w3=which(agg1$Animal=="Elaphurus davidianus" | agg1$Animal=="Ovis aries" | agg1$Animal=="Hippopotamus amphibius" | agg1$Animal=="Giraffa camelopardalis")
w4=which(agg1$Animal=="Equus caballus")
w5=which(agg1$Animal=="Pan troglodytes" | agg1$Animal=="Macaca mulatta")
ord=c(rep(NA,nrow(agg1)))
ord[w1]=1
ord[w2]=2
ord[w3]=3
ord[w4]=4
ord[w5]=5
agg1=cbind(ord,agg1)
colnames(agg1)=c("Order",colnames(agg1)[2:19])
agg1$Order=factor(agg1$Order,levels=c(1:5),labels=c("Odontocetes","Mysticetes","Artiodactyls","Perissodactyls","Primates"))
diet=c(rep(0,210),rep(1,60),rep(2,180),rep(1,60),rep(0,180))
agg1=cbind(diet,agg1)
colnames(agg1)=c("Diet_Binocularity",colnames(agg1)[2:20])
agg1$Diet_Binocularity=factor(agg1$Diet_Binocularity,levels=c(0,1,2),labels=c("A","B","C"))
#A=herbivore-wide field, B=carnivore-lateral, C=omnivore-binocular

agg1=agg1[,c(3:4,1:2,5,6:20)]
row.names(agg1)=Name
rm(Name)
agg1=na.omit(agg1)
agg1.s=na.omit(scale(agg1[,10:19])) #Scaling to avoid distortion due to variables' scales
```

## Global cluster analysis

Each clustering is validated by some appropriate Clustering Validation Indexes (CVI):

-   **Internal indexes**: used to evaluate separation between groups and compactness within groups. Average Silhouette, % of Negative Silhouettes, Calinski-Harabasz Index, Dunn Index (gD33 version, Arbelaitz et al. 2013);

-   **External indexes**: used to evaluate agreement between calculated clusters and a ground truth. Normalized Variation of Information metric, Adjusted Rand Index, Normalized Chi index, P-value of Fisher's association test between clusters and ground truth (H0: absence of association).

For each clustering, a biplot is also shown, in the space of the first two principal components, with points colored according to both calculated clusters and ground truths. Some internal plots from the parameter's selection functions are also visible.

```{r}
crit.2=db.ground.truth(agg1,2,"Cetacean")$CVI
crit.3a=db.ground.truth(agg1,3,"Aquatic")$CVI
crit.3b=db.ground.truth(agg1,3,"Diet_Binocularity")$CVI
crit.5a=db.ground.truth(agg1,5,"Order")$CVI
crit.5b=db.ground.truth(agg1,5,"Shape")$CVI
crit.6=db.ground.truth(agg1,6,"Layer")$CVI
nam=names(crit.3a)
crit.db=matrix(data=c(crit.2,crit.3a,crit.3b,crit.5a,crit.5b,crit.6),ncol=9,byrow = T)
crit.db=cbind(c("db.2","db.3a","db.3b","db.5a","db.5b","db.6"),crit.db)
colnames(crit.db)=c("Clustering",nam)
```

The original data showed much greater agreement between calculated clusters for Shape than for all the other ground truths. In order to uncover patterns hidden from the variability induced by Shape, separate clusterings were calculated for each different level of Shape and for each combination of Shape and cortical Layer. This approach is also shown in this Rmd.

## Ellipsoid

```{r}
s=subset(agg1,agg1$Shape=="Ellipsoid")
crit.2.e=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e=db.ground.truth(s,5,"Order")$CVI
crit.6.e=db.ground.truth(s,6,"Layer")$CVI

crit.db.e=matrix(data=c(crit.2.e,crit.3a.e,crit.3b.e,crit.5.e,crit.6.e),ncol=9,byrow = T)
crit.db.e=cbind(c("db.2.e","db.3a.e","db.3b.e","db.5.e","db.6.e"),crit.db.e)
colnames(crit.db.e)=c("Clustering",nam)
```

### Ellipsoid \* Layer

```{r}
s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==1)
crit.2.e.1=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.1=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.1=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.1=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==2)
crit.2.e.2=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.2=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.2=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.2=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==3)
crit.2.e.3=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.3=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.3=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.3=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==4)
crit.2.e.4=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.4=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.4=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.4=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==5)
crit.2.e.5=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.5=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.5=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.5=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Ellipsoid" & agg1$Layer==6)
crit.2.e.6=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.e.6=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.e.6=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.e.6=db.ground.truth(s,5,"Order")$CVI

crit.db.e.lay=matrix(data=c(crit.2.e.1,crit.3a.e.1,crit.3b.e.1,crit.5.e.1,crit.2.e.2,crit.3a.e.2,crit.3b.e.2,crit.5.e.2,crit.2.e.3,crit.3a.e.3,crit.3b.e.3,crit.5.e.3,crit.2.e.4,crit.3a.e.4,crit.3b.e.4,crit.5.e.4,crit.2.e.5,crit.3a.e.5,crit.3b.e.5,crit.5.e.5,crit.2.e.6,crit.3a.e.6,crit.3b.e.6,crit.5.e.6),ncol=9,byrow = T)
crit.db.e.lay=cbind(c("db.2.e.1","db.3a.e.1","db.3b.e.1","db.5.e.1","db.2.e.2","db.3a.e.2","db.3b.e.2","db.5.e.2","db.2.e.3","db.3a.e.3","db.3b.e.3","db.5.e.3","db.2.e.4","db.3a.e.4","db.3b.e.4","db.5.e.4","db.2.e.5","db.3a.e.5","db.3b.e.5","db.5.e.5","db.2.e.6","db.3a.e.6","db.3b.e.6","db.5.e.6"),crit.db.e.lay)
colnames(crit.db.e.lay)=c("Clustering",nam)
```

## Round

```{r}
s=subset(agg1,agg1$Shape=="Round")
crit.2.r=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r=db.ground.truth(s,5,"Order")$CVI
crit.6.r=db.ground.truth(s,6,"Layer")$CVI

crit.db.r=matrix(data=c(crit.2.r,crit.3a.r,crit.3b.r,crit.5.r,crit.6.r),ncol=9,byrow = T)
crit.db.r=cbind(c("db.2.r","db.3a.r","db.3b.r","db.5.r","db.6.r"),crit.db.r)
colnames(crit.db.r)=c("Clustering",nam)
```

### Round \* Layer

```{r}
s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==1)
crit.2.r.1=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.1=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.1=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.1=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==2)
crit.2.r.2=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.2=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.2=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.2=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==3)
crit.2.r.3=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.3=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.3=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.3=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==4)
crit.2.r.4=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.4=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.4=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.4=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==5)
crit.2.r.5=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.5=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.5=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.5=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Round" & agg1$Layer==6)
crit.2.r.6=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.r.6=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.r.6=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.r.6=db.ground.truth(s,5,"Order")$CVI

crit.db.r.lay=matrix(data=c(crit.2.r.1,crit.3a.r.1,crit.3b.r.1,crit.5.r.1,crit.2.r.2,crit.3a.r.2,crit.3b.r.2,crit.5.r.2,crit.2.r.3,crit.3a.r.3,crit.3b.r.3,crit.5.r.3,crit.2.r.4,crit.3a.r.4,crit.3b.r.4,crit.5.r.4,crit.2.r.5,crit.3a.r.5,crit.3b.r.5,crit.5.r.5,crit.2.r.6,crit.3a.r.6,crit.3b.r.6,crit.5.r.6),ncol=9,byrow = T)
crit.db.r.lay=cbind(c("db.2.r.1","db.3a.r.1","db.3b.r.1","db.5.r.1","db.2.r.2","db.3a.r.2","db.3b.r.2","db.5.r.2","db.2.r.3","db.3a.r.3","db.3b.r.3","db.5.r.3","db.2.r.4","db.3a.r.4","db.3b.r.4","db.5.r.4","db.2.r.5","db.3a.r.5","db.3b.r.5","db.5.r.5","db.2.r.6","db.3a.r.6","db.3b.r.6","db.5.r.6"),crit.db.r.lay)
colnames(crit.db.r.lay)=c("Clustering",nam)
```

## Pyramidal

```{r}
s=subset(agg1,agg1$Shape=="Pyramidal")
crit.2.p=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p=db.ground.truth(s,5,"Order")$CVI
crit.6.p=db.ground.truth(s,6,"Layer")$CVI

crit.db.p=matrix(data=c(crit.2.p,crit.3a.p,crit.3b.p,crit.5.p,crit.6.p),ncol=9,byrow = T)
crit.db.p=cbind(c("db.2.p","db.3a.p","db.3b.p","db.5.p","db.6.p"),crit.db.p)
colnames(crit.db.p)=c("Clustering",nam)
```

### Pyramidal \* Layer

```{r}
s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==1)
crit.2.p.1=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.1=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.1=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.1=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==2)
crit.2.p.2=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.2=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.2=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.2=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==3)
crit.2.p.3=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.3=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.3=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.3=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==4)
crit.2.p.4=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.4=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.4=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.4=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==5)
crit.2.p.5=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.5=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.5=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.5=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Pyramidal" & agg1$Layer==6)
crit.2.p.6=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.p.6=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.p.6=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.p.6=db.ground.truth(s,5,"Order")$CVI

crit.db.p.lay=matrix(data=c(crit.2.p.1,crit.3a.p.1,crit.3b.p.1,crit.5.p.1,crit.2.p.2,crit.3a.p.2,crit.3b.p.2,crit.5.p.2,crit.2.p.3,crit.3a.p.3,crit.3b.p.3,crit.5.p.3,crit.2.p.4,crit.3a.p.4,crit.3b.p.4,crit.5.p.4,crit.2.p.5,crit.3a.p.5,crit.3b.p.5,crit.5.p.5,crit.2.p.6,crit.3a.p.6,crit.3b.p.6,crit.5.p.6),ncol=9,byrow = T)
crit.db.p.lay=cbind(c("db.2.p.1","db.3a.p.1","db.3b.p.1","db.5.p.1","db.2.p.2","db.3a.p.2","db.3b.p.2","db.5.p.2","db.2.p.3","db.3a.p.3","db.3b.p.3","db.5.p.3","db.2.p.4","db.3a.p.4","db.3b.p.4","db.5.p.4","db.2.p.5","db.3a.p.5","db.3b.p.5","db.5.p.5","db.2.p.6","db.3a.p.6","db.3b.p.6","db.5.p.6"),crit.db.p.lay)
colnames(crit.db.p.lay)=c("Clustering",nam)
```

## Complex

```{r}
s=subset(agg1,agg1$Shape=="Complex")
crit.2.c=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c=db.ground.truth(s,5,"Order")$CVI
crit.6.c=db.ground.truth(s,6,"Layer")$CVI

crit.db.c=matrix(data=c(crit.2.c,crit.3a.c,crit.3b.c,crit.5.c,crit.6.c),ncol=9,byrow = T)
crit.db.c=cbind(c("db.2.c","db.3a.c","db.3b.c","db.5.c","db.6.c"),crit.db.c)
colnames(crit.db.c)=c("Clustering",nam)
```

### Complex \* Layer

```{r}
s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==1)
crit.2.c.1=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.1=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.1=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.1=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==2)
crit.2.c.2=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.2=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.2=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.2=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==3)
crit.2.c.3=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.3=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.3=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.3=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==4)
crit.2.c.4=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.4=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.4=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.4=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==5)
crit.2.c.5=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.5=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.5=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.5=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Complex" & agg1$Layer==6)
crit.2.c.6=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.c.6=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.c.6=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.c.6=db.ground.truth(s,5,"Order")$CVI

crit.db.c.lay=matrix(data=c(crit.2.c.1,crit.3a.c.1,crit.3b.c.1,crit.5.c.1,crit.2.c.2,crit.3a.c.2,crit.3b.c.2,crit.5.c.2,crit.2.c.3,crit.3a.c.3,crit.3b.c.3,crit.5.c.3,crit.2.c.4,crit.3a.c.4,crit.3b.c.4,crit.5.c.4,crit.2.c.5,crit.3a.c.5,crit.3b.c.5,crit.5.c.5,crit.2.c.6,crit.3a.c.6,crit.3b.c.6,crit.5.c.6),ncol=9,byrow = T)
crit.db.c.lay=cbind(c("db.2.c.1","db.3a.c.1","db.3b.c.1","db.5.c.1","db.2.c.2","db.3a.c.2","db.3b.c.2","db.5.c.2","db.2.c.3","db.3a.c.3","db.3b.c.3","db.5.c.3","db.2.c.4","db.3a.c.4","db.3b.c.4","db.5.c.4","db.2.c.5","db.3a.c.5","db.3b.c.5","db.5.c.5","db.2.c.6","db.3a.c.6","db.3b.c.6","db.5.c.6"),crit.db.c.lay)
colnames(crit.db.c.lay)=c("Clustering",nam)
```

## Microglia

```{r}
s=subset(agg1,agg1$Shape=="Microglia")
crit.2.m=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m=db.ground.truth(s,5,"Order")$CVI
crit.6.m=db.ground.truth(s,6,"Layer")$CVI

crit.db.m=matrix(data=c(crit.2.m,crit.3a.m,crit.3b.m,crit.5.m,crit.6.m),ncol=9,byrow = T)
crit.db.m=cbind(c("db.2.m","db.3a.m","db.3b.m","db.5.m","db.6.m"),crit.db.m)
colnames(crit.db.m)=c("Clustering",nam)
```

### Microglia \* Layer

```{r}
s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==1)
crit.2.m.1=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.1=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.1=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.1=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==2)
crit.2.m.2=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.2=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.2=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.2=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==3)
crit.2.m.3=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.3=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.3=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.3=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==4)
crit.2.m.4=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.4=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.4=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.4=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==5)
crit.2.m.5=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.5=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.5=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.5=db.ground.truth(s,5,"Order")$CVI

s=subset(agg1,agg1$Shape=="Microglia" & agg1$Layer==6)
crit.2.m.6=db.ground.truth(s,2,"Cetacean")$CVI
crit.3a.m.6=db.ground.truth(s,3,"Aquatic")$CVI
crit.3b.m.6=db.ground.truth(s,3,"Diet_Binocularity")$CVI
crit.5.m.6=db.ground.truth(s,5,"Order")$CVI

crit.db.m.lay=matrix(data=c(crit.2.m.1,crit.3a.m.1,crit.3b.m.1,crit.5.m.1,crit.2.m.2,crit.3a.m.2,crit.3b.m.2,crit.5.m.2,crit.2.m.3,crit.3a.m.3,crit.3b.m.3,crit.5.m.3,crit.2.m.4,crit.3a.m.4,crit.3b.m.4,crit.5.m.4,crit.2.m.5,crit.3a.m.5,crit.3b.m.5,crit.5.m.5,crit.2.m.6,crit.3a.m.6,crit.3b.m.6,crit.5.m.6),ncol=9,byrow = T)
crit.db.m.lay=cbind(c("db.2.m.1","db.3a.m.1","db.3b.m.1","db.5.m.1","db.2.m.2","db.3a.m.2","db.3b.m.2","db.5.m.2","db.2.m.3","db.3a.m.3","db.3b.m.3","db.5.m.3","db.2.m.4","db.3a.m.4","db.3b.m.4","db.5.m.4","db.2.m.5","db.3a.m.5","db.3b.m.5","db.5.m.5","db.2.m.6","db.3a.m.6","db.3b.m.6","db.5.m.6"),crit.db.m.lay)
colnames(crit.db.m.lay)=c("Clustering",nam)
```

## Output

```{r}
crit.db.sha=rbind(crit.db.e,crit.db.r,crit.db.p,crit.db.c,crit.db.m)
crit.db.lay=rbind(crit.db.e.lay,crit.db.r.lay,crit.db.p.lay,crit.db.c.lay,crit.db.m.lay)

write.csv(crit.db,"crit.db.csv")
write.csv(crit.db.sha,"crit.db.sha.csv")
write.csv(crit.db.lay,"crit.db.lay.csv")
```
